This is an example of article title

### 2023-07-08 paper/paper.tex
chore: Paper additions

\section{Overview}
To partition a cloud of points of $d$ dimensions,
\verb|AdaptiveHierarchicalRegularBinning.jl| maps each point to a $1$ dimensional
space filling curve, Morton or Z-order curve. The hierarchical tree is defined by the
Morton index of each point.

@@

  \begin{figure}
    \centerline{
      \includegraphics[width=10pc]{figures/spatial-division.png}
      \includegraphics[width=10pc]{figures/spatial-tree.png}
    }
    \caption{ (a) Spatial division, (b) Spatial tree }
    \label{spatial-tree}
  \end{figure}
\end{example}

\subsection{Reordering}
Sorting in terms of the morton indices groups the points together for efficient access
when the tree is iterated over. We use the \verb|MSD Radixsort| algorithm to perform the
soring.

\subsubsection{MSD Radixsort}
\verb|MSD Radixsort| sorts based on the most significant radix. Afterwards, elements
with the same most significant digit (MSD) form a consecutive sub-array. The algorithm
recurses for each such sub-array selecting the next digit in line. Selecting a large
radix is crucial for the algorithm since a large radix lowers the number of recursions.
\\\\
\verb|MSD Radixsort| respects the hierarchical nature of the target tree and thus it offers
benefits over other sorting algorithms. Each recursion of the algorithm sorts a section
of the resulting tree. If the radix matches the number of dimensions then each recursion
sorts the current level of the tree.
\\\\
When the number of elements of a sub array is small the recursion for that sub-array stops.
This leads to a partialy sorted array by design. Not sorting small sub-arrays leads to less
overhead without sacrificing future performance. Nodes with unsorted morton indices will
always be leaf nodes.
\\\\
\verb|MSD Radixsort| can be trivialy parallelized. After the initial sorting, the formed
sub-arrays are mutually exclusive, meaning that no data dependencies are shared between
them. Each sub-array can then be sorted in parallel.

@@

Our implementation utilizes a watefall scheme to adapt to the needs of the problem at
hand. There are $3$ different thresholds that can tune the runtime performance of the
spatial partitioning.

@@

\subsection{Curse of high dimensionality}
For $n=10^5$ uniformly distributed points we computed the average number of points in a
leaf node of the tree. The graph \ref{b_vs_d} shows that for increasingly higher
dimensional spaces the data get exponatialy sparser.

@@

\subsection{Uniform data 2 dimensions}

\begin{lstlisting}[language = Julia]
julia> d, n = size(X)
(2, 100000)

julia> l = 4;

julia> @time t = regural_bin(UInt128,
                             X,
                             l,
                             1;
                             dims=2);
0.025399 seconds (... 43.93% gc time)
\end{lstlisting}

@@

\begin{figure}[!ht]
  \centerline{
    \includegraphics[width=10pc]{figures/uniform_2d/hist.png}
  }
  \caption{ Distribution of leaf node length }
  \label{uniform-2d-hist}
\end{figure}

@@

\subsection{PBMC 30 dimensions}

@@

  \caption{
    30-Dimensional PBMC data distribution of leaf node length for tree level
    (a) 1 (b) 2 (c) 3 (d) 4.
  }
  \label{pbmc-30-hist}
\end{figure}

\subsection{MNIST 30 dimensions}

Unlike the PBMC dataset, the MNIST dataset does not exhibit significant
outliers. As a result, the distribution between the leaf nodes in the tree
becomes more uniform. Figure \ref{mnist-30-hist} (a) reveals that a majority of
points are concentrated within leaf nodes that possess sufficiently large
lengths. This favorable grouping indicates a more balanced distribution of data
points.
\\\\
However, as we progress to higher levels of the tree, the impact of the
dataset's high dimensionality becomes apparent. Consequently, the points become
exponentially sparser, leading to the formation of singleton leaf nodes.
This observation is clearly demonstrated in Figure \ref{mnist-30-hist} (b).

@@

  \caption{
    30-Dimensional MNIST data distribution of leaf node length for tree level
    (a) 1 (b) 2
  }
  \label{mnist-30-hist}
\end{figure}

@@

  \caption{
    30-Dimensional MNIST data distribution of leaf node length for tree level
    2. Using Small threshold of 10000.
  }
  \label{mnist-30-hist-th}
\end{figure}

% TODO Remove
\pagebreak
\section{Juliacon stuff, kept for reference}
\section{Introduction}

The \LaTeX{} document preparation system is a special version of the
\TeX; typesetting program where-in a collection of \TeX; commands
are added to \LaTeX{} to simplify typesetting. Importantly, it allows
the author to concentrate on the logical structure of the document
rather than its visual layout.\vskip 6pt
Moreover, \LaTeX{} provides a consistent and comprehensive document
preparation interface. There are simple-to-use commands for
generating a table of contents, lists of figures and/or tables, and indexes.
\LaTeX{} can automatically number list entries, equations, figures,
tables, and footnotes, as well as articles, sections, and subsections.
Using this numbering system, bibliographic citations, page
references, and cross references to any other numbered entity (e.g.
article, section, equation, figure, list entry, etc.) become quite simple
and straightforward. The use of \LaTeX{} document classes allows
a simple change of class to transform the appearance of your document.\vskip 6pt
\LaTeX{} is a powerful tool for managing long and complex documents.
In particular, partial processing enables long documents to
be produced article by article without losing sequential information.
The use of document classes allows a simple change of style
(or style option) to transform the appearance of your document.

\section{The JuliaCon Article Class}
\label{sec:documentclass}
%
The juliacon class file preserves the standard LATEX{} interface such
that any document that can be produced using the standard LATEX{}
article class can also be produced with the class file.\vskip 6pt
It is likely that the make up will change after file submission. For
this reason, we ask you to ignore details such as slightly long lines,
page stretching, or figures falling out of synchronization, as these
details can be dealt with at a later stage.\vskip 6pt
Use should be made of symbolic references (\verb|\ref| ) in order to
protect against late changes of order, etc.

\section{USING THE JuliaCon Article CLASS FILE}

If the file \verb|juliacon.cls|  is not already in the appropriate system directory
for \LaTeX{} files, either arrange for it to be put there or copy
it to your working directory. The \verb|juliacon|  document class is implemented
as a complete class, not a document style option. In order to
use the \verb|juliacon  document class, replace \verb|article|  by \verb|juliacon|  in the
\verb|\documentclass|  command at the beginning of your document:
\vskip 6pt
\begin{centering}
    \verb|\documentclass{article}|  \end{centering}
\vskip 6pt
replace by
\vskip 6pt
 \verb|\documentclass{juliacon}|  \vskip 6pt
In general, the following standard document \verb|style|  options should
{ \itshape not} be used with the {\footnotesize \itshape article} class file:
\begin{enumerate}
\item[(1)] \verb|10pt|,  \verb|11pt|,  \verb|12pt|   ? unavailable;
\item[(2)] \verb|twoside|  (no associated style file) ? \verb|twoside|  is the default;
\item[(3)] \verb|fleqn|, \verb|leqno|, \verb|titlepage| ? should not be used;
\end{enumerate}

\section{Additional Document Style Options}
\label{sec:additional_doc}
%
The following additional style option is available with the \verb|juliacon|  class file:
\vskip 6pt
Please place any additional command definitions at the very start of
the \LaTeX{} file, before the \verb|\begin{document}|. For example, user-defined
\verb|\def|  and \verb|\newcommand|   commands that define macros for
technical expressions should be placed here. Other author-defined
macros should be kept to a minimum.
\vskip 6pt
Commands that differ from the standard \LaTeX{} interface, or that
are provided in addition to the standard interface, are explained in
this guide. This guide is not a substitute for the \LaTeX{} manual itself.
Authors planning to submit their papers in \LaTeX{} are advised to use
\verb|juliacon.cls|  as early as possible in the creation of their files.

%
%
%
%
\begin{table*}[t]
\tabcolsep22pt
\tbl{If necessary, the tables can be extended both columns.}{
\begin{tabular}{|l|l|c|c|}\hline
Label & \multicolumn{1}{c|}{Description}
& Number of Users &
Number of Queries\\\hline
Test 1 & Training Data &
\smash{\raise-7pt\hbox{70}} & 104\\
\cline{1-2}\cline{4-4}
Test 2 & Testing Data I & & 105\\\hline
Test 3 & Testing Data II & 30 & 119\\\hline
& Total & 100 & 328\\\hline
\end{tabular}}
\label{tab:symbols}
\begin{tabnote}
This is an example of table footnote.
\end{tabnote}
\end{table*}
% \begin{figure*}[t]
% \centerline{\includegraphics[width=11cm]{juliagraphs.png}}
% \caption{If necessary, the images can be extended both columns.}
%   \label{fig:sample_image}
% \end{figure*}

\section{Additional features}
\label{sec:additional_faci}
In addition to all the standard \LaTeX{} design elements, the \verb|juliacon|  class file includes the following features:
In general, once you have used the additional \verb|juliacon.cls| facilities
in your document, do not process it with a standard \LaTeX{} class
file.

\subsection{Titles, Author's Name, and Affiliation}
\label{subsub:title_auth}
The title of the article, author's name, and affiliation are used at the
beginning of the article (for the main title). These can be produced
using the following code:

\begin{verbatim}
\title{ This is an example of article title} }
\author{
   \large 1st Author \\[-3pt]
   \normalsize 1st author's affiliation  \\[-3pt]
    \normalsize 1st line of address \\[-3pt]
    \normalsize 2nd line of address \\[-3pt]
    \normalsize	1st author's email address \\[-3pt]
  \and
   \large 2nd Author \\[-3pt]
   \normalsize 2nd author's affiliation  \\[-3pt]
    \normalsize 1st line of address \\[-3pt]
    \normalsize 2nd line of address \\[-3pt]
    \normalsize	2nd author's email address \\[-3pt]
\and
   \large 3rd Author \\[-3pt]
   \normalsize 3rd author's affiliation  \\[-3pt]
    \normalsize 1st line of address \\[-3pt]
    \normalsize 2nd line of address \\[-3pt]
    \normalsize	3rd author's email address \\[-3pt]
}
\maketitle
\end{verbatim}

\subsection{Writing Julia code}

A special environment is already defined for Julia code,
built on top of \textit{listings} and \textit{jlcode}.

\begin{verbatim}
\begin{lstlisting}[language = Julia]
using Plots

x = -3.0:0.01:3.0
y = rand(length(x))
plot(x, y)
\end{lstlisting}
\end{verbatim}
\begin{lstlisting}[language = Julia]
using Plots

x = -3.0:0.01:3.0
y = rand(length(x))
plot(x, y)
\end{lstlisting}

\subsection{Abstracts, Key words, term etc...}
\label{subsub:abs_key_etc}

At the beginning of your article, the title should be generated
in the usual way using the \verb|\maketitle|  command. For genaral tem and keywords use
\verb|\terms|,
\verb|\keywords|  commands respectively. The abstract should be enclosed
within an abstract environment, All these environment
can be produced using the following code:
\begin{verbatim}
\terms{Experimentation, Human Factors}

\keywords{Face animation, image-based modelling...}

\begin{abstract}
In this paper, we propose a new method for the
systematic determination of the model's base of
time varying delay system. This method based on
the construction of the classification data related
to the considered system. The number, the orders,
the time delay and the parameters of the local
models are generated automatically without any
knowledge about the full operating range of the
process. The parametric identification of the local
models is realized by a new recursive algorithm for
on line identification of systems with unknown time
delay. The proposed algorithm allows simultaneous
estimation of time delay and parameters of
discrete-time systems. The effectiveness of
the new method has been illustrated through
simulation.
\end{abstract}

\end{verbatim}

\section{Some guidelines}
\label{sec:some_guide}
The following notes may help you achieve the best effects with the
\verb|juliacon|  class file.

\subsection{Sections}
\label{subsub:sections}
\LaTeXe{}  provides four levels of section headings and they are all
defined in the \verb|juliacon|  class file:
\begin{itemize}
\item \verb|\section|
\item \verb|\subsection|
\item \verb|\subsubsection|
\item \verb|\paragraph|
\end{itemize}
Section headings are automatically converted to allcaps style.
\subsection{Lists}
\label{sec:lists}
%
The \verb|juliacon|   class file provides unnumbered lists using the
\verb|unnumlist|  environment for example,

\begin{unnumlist}
\item First unnumbered item which has no label and is indented from the
left margin.
\item Second unnumbered item.
\item Third unnumbered item.
\end{unnumlist}
The unnumbered list which has no label and is indented from the
left margin. was produced by:
\begin{verbatim}
\begin{unnumlist}
\item First unnumbered item...
\item Second unnumbered item...
\item Third unnumbered item...
\end{unnumlist}
\end{verbatim}

The \verb|juliacon|  class file also provides hyphen list using the
\verb|itemize|  environment for example,
\begin{itemize}
\item First unnumbered bulleted item which has no label and is indented
from the left margin.
\item Second unnumbered bulleted item.
\item Third unnumbered bulleted item which has no label and is indented
from the left margin.
\end{itemize}
was produced by:
\begin{verbatim}
\begin{itemize}
\item First item...
\item Second item...
\item Third item...
\end{itemize}
\end{verbatim}

Numbered list is also provided in acmtog class file using the
enumerate environment for example,
\begin{enumerate}
\item The attenuated and diluted stellar radiation.
\item Scattered radiation, and
\item Reradiation from other grains.
\end{enumerate}

was produced by:
\begin{verbatim}
\begin{enumerate}
\item The attenuated...
\item Scattered radiation, and...
\item Reradiation from other grains...
\end{enumerate}
\end{verbatim}
\subsection{Illustrations (or figures)}
\label{subsub:sec_Illus}
The \verb|juliacon|  class file will cope with most of the positioning of
your illustrations and you should not normally use the optional positional
qualifiers on the \verb|figure|  environment that would override
these decisions.
\vskip 6pt

%
\begin{figure}[t]
\centerline{\includegraphics[width=4cm]{juliagraphs.png}}
\caption{This is example of the image in a column.}
	\label{fig:sample_figure}
\end{figure}

The figure \ref{fig:sample_figure} is taken from the JuliaGraphs
organization \footnote{https://github.com/JuliaGraphs}.

Figure captions should be \emph{below} the figure itself, therefore the
\verb|\caption|  command should appear after the figure or space left for
an illustration. For example, Figure 1 is produced using the following
commands:

\begin{verbatim}
\begin{figure}
\centerline{\includegraphics[width=20pc]{Graphics.eps}}
\caption{An example of the testing process for a
binary tree. The globa null hypothesis is tested
first at level $\alpha$ (a), and the level of
individual variables is reached last (d). Note
that individual hypotheses can be tested at
level $\alpha/4$ and not $\alpha/8$ as one might
expect at first.}
\label{sample-figure_2}
\end{figure}
\end{verbatim}
Figures can be resized using first and second argument of
\verb|\includegraphics|   command. First argument is used for modifying
figure height and the second argument is used for modifying
figure width respectively.
\vskip 6pt
Cross-referencing of figures, tables, and numbered, displayed
equations using the \verb|\label|  and \verb|\ref|   commands is encouraged.
For example, in referencing Figure 1 above, we used
\verb|Figure~\ref{sample-figure}|

 \subsection{Tables}
\label{subsub:sec_Tab}
The \verb|juliacon|   class file will cope with most of the positioning of
your tables and you should not normally use the optional positional qualifiers on the table environment which would override these
decisions. Table captions should be at the top.
\begin{verbatim}
\begin{table}
\tbl{Tuning Set and Testing Set}{
\begin{tabular}{|l|l|c|c|}\hline
Label & \multicolumn{1}{c|}{Description}
& Number of Users &
Number of Queries\\\hline
Train70 & Training Data &
\smash{\raise-7pt\hbox{70}} & 104\\
\cline{1-2}\cline{4-4}
Test70 & Testing Data I & & 105\\\hline
Test30 & Testing Data II & 30 & 119\\\hline
& Total & 100 & 328\\\hline
\end{tabular}}
\end{table}
\end{verbatim}

\begin{table}
\tbl{Tuning Set and Testing Set}{
\begin{tabular}{|l|l|c|c|}\hline
Label & \multicolumn{1}{c|}{Description}
& Number of Users &
Number of Queries\\\hline
Test 1 & Training Data &
\smash{\raise-7pt\hbox{70}} & 104\\
\cline{1-2}\cline{4-4}
Test 2 & Testing Data I & & 105\\\hline
Test 3 & Testing Data II & 30 & 119\\\hline
& Total & 100 & 328\\\hline
\end{tabular}}
\end{table}
\subsection{Landscaping Pages}
\label{subsub:landscaping_pages}
If a table is too wide to fit the standard measure, it may be turned,
with its caption, to 90 degrees. Landscape tables cannot be produced
directly using the \verb|juliacon|   class file because \TeX{} itself cannot
turn the page, and not all device drivers provide such a facility.
The following procedure can be used to produce such pages.
\vskip 6pt
Use the package \verb|rotating|   in your document and change the coding
from
\begin{verbatim}
\begin{table}...\end{table}
to
\begin{sidewaystable}...\end{sidewaystable}
and for figures
\begin{figure}...\end{figure}
to
\begin{sidewaysfigure}...\end{sidewaysfigure}
\end{verbatim}

environments in your document to turn your table on the appropriate
page of your document. For instance, the following code prints
a page with the running head, a message half way down and the
table number towards the bottom.
\begin{verbatim}
\begin{sidewaystable}
\tbl{Landscape table caption to go here.}{...}
\label{landtab}
\end{sidewaystable}
\end{verbatim}

\subsection{Double Column Figure and Tables}
\label{subsub:double_fig_tab}
For generating the output of figures and tables in double column
we can use the following coding:

\begin{enumerate}
\item For Figures:
\begin{verbatim}
\begin{figure*}...\end{figure*}
\end{verbatim}
\item For landscape figures:
\begin{verbatim}
\begin{sidewaysfigure*}...\end{sidewaysfigure*}
\end{verbatim}
\item For Tables:
\begin{verbatim}
\begin{table*}...\end{table*}
\end{verbatim}
\item For landscape tables:
\begin{verbatim}
\begin{sidewaystable*}...\end{sidewaystable*}
\end{verbatim}
\end{enumerate}

\subsection{Typesetting Mathematics}
\label{subsub:type_math}
The \verb|juliacon| class file will set displayed mathematics with center to
the column width, provided that you use the \LaTeXe{} standard of
open and closed square brackets as delimiters.
The equation
\[
\sum_{i=1}^p \lambda_i = (S)
\]

was typeset using the acmtog class file with the commands

\begin{verbatim}
\[
\sum_{i=1}^p \lambda_i = (S)
\]
\end{verbatim}

For display equations, cross-referencing is encouraged. For example,
\begin{verbatim}
\begin{equation}
(n-1)^{-1} \sum^n_{i=1} (X_i - \overline{X})^2.
\label{eq:samplevar}
\end{equation}
Equation~(\ref{eq:samplevar}) gives the formula for
sample variance.
\end{verbatim}
The following output is generated with the above coding:
\begin{equation}
(n-1)^{-1} \sum^n_{i=1} (X_i - \overline{X})^2.
\label{eq:samplevar}
\end{equation}
Equation~(\ref{eq:samplevar}) gives the formula for
sample variance.

\subsection{Enunciations}
\label{subsub:enunciation}
The \verb|juliacon|   class file generates the enunciations with the help of
the following commands:
\begin{verbatim}
\begin{theorem}...\end{theorem}
\begin{strategy}...\end{strategy}
\begin{property}...\end{property}
\begin{proposition}...\end{proposition}
\begin{lemma}...\end{lemma}
\begin{example}...\end{example}
\begin{proof}...\end{proof}
\begin{definition}...\end{definition}
\begin{algorithm}...\end{algorithm}
\begin{remark}...\end{remark}
\end{verbatim}
The above-mentioned coding can also include optional arguments
such as
\begin{verbatim}
\begin{theorem}[...]. Example for theorem:
\begin{theorem}[Generalized Poincare Conjecture]
Four score and seven ... created equal.
\end{theorem}
\end{verbatim}

\begin{theorem}[Generalized Poincare Conjecture]
Four score and seven years ago our fathers brought forth,
upon this continent, a new nation, conceived in Liberty,
 and dedicated to the proposition that all men are
created equal.
\end{theorem}

\subsection{Extract}
\label{subsub:extract}
Extract environment should be coded within
\begin{verbatim}
\begin{extract}..\end{extract}
\end{verbatim}

\subsection{Balancing column at last page}
\label{subsub:Balance}
For balancing the both column length at last page use :
\begin{verbatim}
\vadjust{\vfill\pagebreak}
\end{verbatim}

%\vadjust{\vfill\pagebreak}

at appropriate place in your \TeX{} file or in bibliography file.

\section{Handling references}
\label{subsub:references}
References are most easily (and correctly) generated using the
BIBTEX, which is easily invoked via
\begin{verbatim}
\bibliographystyle{juliacon}
\bibliography{ref}
\end{verbatim}
When submitting the document source (.tex) file to external
parties, the ref.bib file should be sent with it.
\cite{bezanson2017julia}

>>>>

\section{The need for space partitioning algorithms}
Space partitioning algorithms play a pivotal role in improving the runtime
performance of computational algorithms by dividing data into smaller,
manageable sections. These algorithms are designed to optimize the performance
of computational algorithms by efficiently partitioning the data space in a
manner that aligns with the specific requirements of the problem domain.
\\\\
In this section, we delve into the need for specific space partitioning
algorithms and their significance in addressing the computational bottlenecks
faced by certain algorithms. We examine three prominent application domains:
k-nearest neighbors (kNN) search, collision detection, and the fast multipole
method (FMM). These domains represent scenarios where the all-vs-all model,
involving computations between every pair of data points, hampers the runtime
performance of algorithms.

\subsection{kNN}
One common challenge encountered in computational algorithms, such as k-nearest
neighbors (kNN) search, is the need to compute distances between each pair of
points in the data space. This all-vs-all model becomes computationally
expensive as the dataset size increases. However, by employing a space
partitioning algorithm, the number of distance computations can be
significantly reduced. Instead of calculating distances between all points,
the algorithm subdivides the space into smaller sections, enabling the
identification of nearest neighbors within localized partitions. This approach
drastically improves the runtime performance of the kNN algorithm by reducing
the computational burden associated with distance calculations.

\subsection{Collision Detection}
Collision detection algorithms also stand to benefit from specific space
partitioning techniques. In scenarios where it is necessary to determine if a
collision occurs between objects in a three-dimensional space, the all-vs-all
model becomes impractical for large-scale simulations. By partitioning the
space into smaller regions or bins, collision detection algorithms can focus
computations solely on objects within the same partition. This significantly
reduces the number of pairwise collision checks required and provides a
substantial performance improvement in terms of runtime. Specific space
partitioning algorithms enable efficient identification of potential collision
candidates, contributing to faster and more accurate collision detection in
simulations and virtual environments.

\subsection{FMM}

The fast multipole method (FMM), explored in \cite{FMM} and \cite{AMM}, is an
important algorithm used for solving the N-body problem, which involves the
interaction of multiple particles in a system. The traditional implementation
of FMM suffers from a high computational complexity, as it requires pairwise
computations between all particles in the system. However, space partitioning
algorithms offer an opportunity to optimize the FMM by dividing the space into
subregions or bins. By employing space partitioning techniques, the FMM
algorithm can efficiently identify distant interactions, reducing the number of
pairwise computations required and significantly improving the overall runtime
performance. The incorporation of space partitioning algorithms in FMM enables
the efficient solution of the N-body problem and opens doors for the simulation
of large-scale physical systems.

\section{Implementation}
The partitioning of a $d$-dimensionsional cloud of points is a fundamental task
in many computational applications. The
\verb|AdaptiveHierarchicalRegularBinning.jl|
algorithm offers an effective approach to address this challenge. By utilizing
the Morton curve, a type of one-dimensional space filling curve, this algorithm
maps each point from the $d$-dimensional space to a one-dimensional
representation. Consequently, the hierarchical tree structure is established
based on the Morton index assigned to each point.
\\\\
The \verb|AdaptiveHierarchicalRegularBinning.jl| algorithm leverages the
inherent properties of the Morton curve to efficiently partition the cloud of
points. The Morton curve exhibits a desirable characteristic of spatial
locality preservation, where nearby points in the original $d$-dimensional
space tend to have adjacent Morton indices in the one-dimensional space. This
property ensures that points with similar spatial coordinates are grouped
together in the resulting hierarchical tree structure.

@@

  \begin{figure}[!ht]
    \centerline{
      \includegraphics[width=10pc]{figures/spatial-division.png}
      \includegraphics[width=10pc]{figures/spatial-tree.png}
    }
    \caption{ (a) Spatial division, (b) Spatial tree }
    \label{spatial-tree}
  \end{figure}
\end{example}

\subsection{Reordering}
Sorting the points based on their Morton indices is a key operation in the
\verb|AdaptiveHierarchicalRegularBinning.jl| algorithm, enabling efficient
access during tree iteration. To achieve this, the algorithm employs the
\verb|MSD Radixsort| \cite{radixsort} algorithm, or Most Significant Digit
Radixsort, which examines the most significant bits of the Morton indices. By
grouping points with similar prefixes together, the sorted order achieved
through \verb|MSD Radixsort| enhances data locality and improves performance.
This arrangement ensures that neighboring points in the original $d$-dimensional
space are adjacent in the sorted sequence, optimizing point retrieval and
overall algorithm efficiency.

\subsubsection{MSD Radixsort}
\verb|MSD Radixsort| sorts based on the most significant radix. Afterwards, elements
with the same most significant digit (MSD) form a consecutive sub-array. The algorithm
recurses for each such sub-array selecting the next digit in line, leading to
$O(n\cdot m)$ runtime complexity, where $n$ is the size of the array and $m$ is
the number of recursions. Selecting a large radix is crucial for the algorithm
since a large radix lowers the number of recursions.
\\\\
\verb|MSD Radixsort| respects the hierarchical nature of the target tree and thus it offers
benefits over other sorting algorithms. Each recursion of the algorithm sorts a section
of the resulting tree. If the radix matches the number of dimensions then each recursion
sorts the current level of the tree.
\\\\
When the number of elements of a sub array is small the recursion for that sub-array stops.
This leads to a partialy sorted array by design. Not sorting small sub-arrays leads to less
overhead without sacrificing future performance. Nodes with unsorted morton indices will
always be leaf nodes.
\\\\
\verb|MSD Radixsort| can be trivialy parallelized. After the initial sorting, the formed
sub-arrays are mutually exclusive, meaning that no data dependencies are shared between
them. Each sub-array can then be sorted in parallel.

@@

Our implementation utilizes a watefall scheme to adapt to the needs of the problem at
hand. There are three different thresholds that can tune the runtime performance of the
spatial partitioning.

@@

@@

\subsection{Point per node distribution}

\subsubsection{Uniform data 2 dimensions}

@@

\begin{figure}[!ht]
  \centerline{
    \includegraphics[width=10pc]{figures/experiments/uniform_2d/hist.png}
  }
  \caption{ Distribution of leaf node length }
  \label{uniform-2d-hist}
\end{figure}

@@

\subsubsection{Curse of high dimensionality}
For $n=10^5$ uniformly distributed points we computed the average number of points in a
leaf node of the tree. The graph \ref{b_vs_d} shows that for increasingly higher
dimensional spaces the data get exponatialy sparser.

\subsubsection{PBMC 30 dimensions}

@@

  \caption{
    30-Dimensional PBMC data distribution of leaf node length per leaf for tree
    level (a) 1 (b) 2 (c) 3 (d) 4.
  }
  \label{pbmc-30-hist}
\end{figure}

\subsubsection{MNIST 30 dimensions}

Unlike the PBMC dataset, the MNIST dataset does not exhibit significant
outliers. As a result, the distribution between the leaf nodes in the tree
becomes more uniform. Figure \ref{mnist-30-hist} (a) reveals that a majority of
points are concentrated within leaf nodes that possess sufficiently large
lengths. This favorable grouping indicates a more balanced distribution of data
points.
\\\\
However, as we progress to lower levels of the tree, the impact of the
dataset's high dimensionality becomes apparent. Consequently, the points become
exponentially sparser, leading to the formation of singleton leaf nodes.
This observation is clearly demonstrated in Figure \ref{mnist-30-hist} (b).

@@

  \caption{
    30-Dimensional MNIST data distribution of leaf node length per point for
    tree level (a) 1 (b) 2
  }
  \label{mnist-30-hist}
\end{figure}

@@

  \caption{
    30-Dimensional MNIST data distribution of leaf node length per point for
    tree of max level 2. Using Small threshold of 10000.
  }
  \label{mnist-30-hist-th}
\end{figure}
